02_DESeq_characterization

### 2019-07-24 analysis/03_DESeq2_characterization.Rmd
first pass usical notes

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)

@@

## Manipulation data

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
m.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
m.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# set levels
levels(m.colData$treatment)
m.colData$treatment <- factor(m.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))

m.colData$sextissue <- as.factor(paste(m.colData$sex, m.colData$tissue, sep = "_"))
summary(m.colData[c(7,3,4,5,8)])
```

@@

```{r dds, cache = T}
dds.female_hypothalamus <- subsetDESeq("female_hypothalamus")
dds.female_pituitary <- subsetDESeq("female_pituitary" )
dds.female_gonad <- subsetDESeq("female_gonad" )
dds.male_hypothalamus <- subsetDESeq("male_hypothalamus" )
dds.male_pituitary <- subsetDESeq("male_pituitary"  )
dds.male_gondad <- subsetDESeq("male_gonad")
```

@@

```{r pheatmap, eval = T}
makepheatmap(dds.female_hypothalamus, "female hypothalamus")
makepheatmap(dds.female_pituitary, "female pituitary")
makepheatmap(dds.female_gonad, "female gonad")
makepheatmap(dds.male_hypothalamus, "male hypothalamus")
makepheatmap(dds.male_pituitary, "male pituitary")
makepheatmap(dds.male_gondad, "male gonad")
```
## plot candidate genes

```{r candidates}
plotcandidates(dds.female_hypothalamus, "female hypothalamus")
plotcandidates(dds.female_pituitary, "female pituitary")
plotcandidates(dds.female_gonad, "female gonad")
plotcandidates(dds.male_hypothalamus, "male hypothalamus")
plotcandidates(dds.male_pituitary, "male pituitary")
plotcandidates(dds.male_gondad, "male gonad")
```

>>>>

```{r setup, message=F, comment=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(DESeq2)
library(cowplot)
library(RColorBrewer)
library(pheatmap)
library(kableExtra)
library(viridis)
library("ggimage")

@@

## Characterization data

```{r readfiles}
# import "colData" which contains sample information and "countData" which contains read counts
c.colData <- read.csv("../metadata/00_colData_characterization.csv", header = T, row.names = 1)
c.countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
geneinfo <- read.csv("../metadata/00_geneinfo.csv", row.names = 1)

# set levels
c.colData$treatment <- factor(c.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
levels(c.colData$treatment)
c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
summary(c.colData[c(7,3,4,5,8)])
```

## plot raw counts

```{r musicalgenes, cache = T, eval=T}
head(countData)
head(geneinfo)

candidataCounts <- countData
candidataCounts$Name <- row.names(candidataCounts)
candidataCounts <- candidataCounts %>%
  filter(Name %in% c("NP_990797.2", "NP_001138861.1"))
row.names(candidataCounts) <- candidataCounts$Name
candidataCounts$Name <- NULL
candidataCountsT <- t(candidataCounts)
candidataCountsT <- as.data.frame(candidataCountsT)

candidataCountsT$entrez <- row.names(candidataCountsT)
colData$entrez <- row.names(colData)

candidataCountsTcol <- full_join(colData, candidataCountsT)

str(candidataCountsTcol)

candidataCountsTcol2 <- candidataCountsTcol %>%
  group_by(sextissue, treatment) %>%
  summarize(GAL = mean(NP_001138861.1, na.rm = TRUE),
            PRL = mean(NP_990797.2, na.rm = TRUE)) %>%
  gather(GAL, PRL, key = "gene", value = "mean")

candidataCountsTcol2 <- as.data.frame(candidataCountsTcol2)

musicgenes <- function(mygroup, mysusbtitle){

myplot <- candidataCountsTcol2 %>%
  filter(sextissue %in% c(mygroup)) %>%
ggplot(aes(x = treatment, y = mean)) +
  geom_point(colour = "white") + 
  geom_line(group = 1, color='grey') +
  theme_bw() +
  facet_wrap(~gene, scales = "free", ncol = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_image(aes(image="https://storage.needpix.com/rsynced_images/music-34177_1280.png")) +
  labs(y = "mean gene expression", subtitle = mysusbtitle, x = NULL)

  return(myplot)

}

a <- musicgenes("female_hypothalamus", "female hypothalamus")
b <- musicgenes("male_hypothalamus", "male hypothalamus")

c <- plot_grid(a,b)
c

pdf("../figures/characterization/musicalgenes-1.pdf", width = 8, height = 6)
plot(c)
dev.off()

```

@@

```{r dds, cache = T}
dds.female_hypothalamus <- subsetDESeq(c.colData,  c.countData, "female_hypothalamus")
dds.female_pituitary <- subsetDESeq(c.colData,  c.countData, "female_pituitary" )
dds.female_gonad <- subsetDESeq(c.colData,  c.countData, "female_gonad" )
dds.male_hypothalamus <- subsetDESeq(c.colData,  c.countData, "male_hypothalamus" )
dds.male_pituitary <- subsetDESeq(c.colData,  c.countData, "male_pituitary"  )
dds.male_gondad <- subsetDESeq(c.colData,  c.countData, "male_gonad")
```

@@

```{r pheatmap, eval = F}
makepheatmap(dds.female_hypothalamus, "female hypothalamus")
makepheatmap(dds.female_pituitary, "female pituitary")
makepheatmap(dds.female_gonad, "female gonad")
makepheatmap(dds.male_hypothalamus, "male hypothalamus")
makepheatmap(dds.male_pituitary, "male pituitary")
makepheatmap(dds.male_gondad, "male gonad")
```
## plot candidate genes

```{r candidates, eval = F}
plotcandidates(dds.female_hypothalamus, "female hypothalamus")
plotcandidates(dds.female_pituitary, "female pituitary")
plotcandidates(dds.female_gonad, "female gonad")
plotcandidates(dds.male_hypothalamus, "male hypothalamus")
plotcandidates(dds.male_pituitary, "male pituitary")
plotcandidates(dds.male_gondad, "male gonad")
```
